### 2017-02-20 abstract.tex
preliminary,data-independence,rules

\begin{abstract}
I am abstract
\end{abstract}

\forget{
\noindent Keywords: weak memory model, $\textit{linearizability}$,
$\textit{TSO-to-TSO linearizability}$
}