### 2017-02-20 data-independence-PQ.tex

\section{Data-Independence of Priority Queue}
\label{sec:data-independence of priority queue}

Data-independence \cite{Wolper:1986} can be used to effectively handle unbounded data domain. In this section, we slightly modify the notion of data-independence in \cite{Wolper:1986} and propose data-differentiated sequences and data-independence for priority queues.

Let $\_$ denote an element, of which the value is irrelevant. A sequential execution $e$ of priority is said to be data-differentiated if, for all $d \in \mathbb{D}$, there is at most one method event $\textit{put}(d,\_)$ in $e$. Note that a data-differentiated sequential execution $e$ may contains more than one items with a same priority. The subset of data-differentiated sequential executions of a set $S$ is denoted by $S_{\neq}$. The definition extends to (sets of) executions and histories. For instance, an execution is data-differentiated if, for all $d \in \mathbb{D}$, there is at most one $\textit{cal}_{\_}(\textit{put},d,\_)$.

\begin{example}\label{example:data-differentiated}
$\textit{cal}_{o_1}(\textit{put},a,7) \cdot \textit{ret}_{o_1}(\textit{put},a) \cdot \textit{cal}_{o_2}(\textit{put},a,8) \cdot \textit{ret}_{o_2}(\textit{put},a)$ is not data-differentiated, since there are two $\textit{put}$ methods with the same item.
\end{example}

A renaming function $r$ for priority queue is a function from $\mathbb{D}$ to $\mathbb{D}$. Given a sequential execution (resp., execution or history) $e$, we denote by $r(e)$ the sequential execution (resp., execution or history) obtained from $e$ by replacing every item $x$ by $r(x)$. Note that here the renaming functions rename only the items and keep the priorities unchanged.

%\vspace{-6pt}
\begin{definition}\label{def:priority-value data-independence}
The set of sequential executions (resp., executions or histories) $S$ is data-independent, if:
\begin{itemize}
\setlength{\itemsep}{0.5pt}
\item[-] for all $e \in S$, there exists $e' \in S'$, and a renaming function $r$, such that $e=r(e')$,

\item[-] for all $e \in S$ and for all renaming $r$, $r(e) \in S$.
\end{itemize}
\end{definition}

The following lemma states that, when checking that a data-independent implementation $\mathcal{I}$ is linearizable with respect to a data-independent specification, it is enough to do so for data-differentiated executions, similar as that in \cite{Abdulla:2013}, where the notion of data-independence and differentiated in \cite{Wolper:1986} is used. Thus, in the remainder of the paper, we focus on characterizing linearizability for data-differentiated executions, rather than arbitrary ones. The proof of this lemma can be found in Appendix \ref{sec:appendix in section data-independence of PQ}.

\begin{restatable}{lemma}{DataDifferentiatedisEnoughforPQ}
\label{lemma:data differentiated is enough for PQ}
A data-independent implementation $\mathcal{I}$ is linearizable with respect to a data-independent specification $S$, if and only if $\mathcal{I}_{\neq}$ is linearizable with respect to $S_{\neq}$.
\end{restatable}