### 2017-04-04 data-independence-EPQ.tex

\section{Data-Independence of Extended Priority Queue}
\label{sec:data-independence of extended priority queue}

Data-independence \cite{Wolper:1986} can be used to effectively handle unbounded data domain. In this section, we slightly modify the notion of data-independence in \cite{Wolper:1986} and propose data-differentiated sequences and data-independence for extended priority queues.

Let $\_$ denote an element, of which the value is irrelevant. A sequential execution $e$ of extended priority queue is said to be data-differentiated if, for all $d \in \mathbb{D}$, there is at most one method event $\textit{put}(d,\_)$ in $e$. Note that a data-differentiated sequential execution $e$ may contains more than one items with a same priority. The subset of data-differentiated sequential executions of a set $S$ is denoted by $S_{\neq}$. The definition extends to (sets of) executions and histories. For instance, an execution is data-differentiated if, for all $d \in \mathbb{D}$, there is at most one $\textit{cal}_{\_}(\textit{put},d,\_)$.

\begin{example}\label{example:data-differentiated}
$\textit{cal}_{o_1}(\textit{put},a,7) \cdot \textit{ret}_{o_1}(\textit{put},a) \cdot \textit{cal}_{o_2}(\textit{put},a,8) \cdot \textit{ret}_{o_2}(\textit{put},a)$ is not data-differentiated, since there are two $\textit{put}$ methods with the same item.
\end{example}

A renaming function $r$ for extended priority queue is a function from $\mathbb{D}$ to $\mathbb{D}$. Given a sequential execution (resp., execution or history) $e$, we denote by $r(e)$ the sequential execution (resp., execution or history) obtained from $e$ by replacing every item $x$ by $r(x)$. Note that here the renaming functions rename only the items and keep the priorities unchanged. The intuitive explanation is that renaming items will not influence the executions of program, while renaming priorities may influence the executions of program.

%\vspace{-6pt}
\begin{definition}\label{def:priority-value data-independence}
A set of sequential executions (resp., executions or histories) $S$ is data-independent, if:
\begin{itemize}
\setlength{\itemsep}{0.5pt}
\item[-] for all $e \in S$, there exists $e' \in S'$, and a renaming function $r$, such that $e=r(e')$,

\item[-] for all $e \in S$ and for all renaming $r$, $r(e) \in S$.
\end{itemize}
\end{definition}

The following lemma states that, when checking that a data-independent implementation $\mathcal{I}$ is linearizable with respect to a data-independent specification, it is enough to do so for data-differentiated executions, similar as that in \cite{Abdulla:2013}, where the notion of data-independence and differentiated in \cite{Wolper:1986} is used. Thus, in the remainder of the paper, we focus on characterizing linearizability for data-differentiated executions, rather than arbitrary ones. The proof of this lemma can be found in Appendix \ref{sec:appendix in section data-independence of EPQ}.

\begin{restatable}{lemma}{DataDifferentiatedisEnoughforPQ}
\label{lemma:data differentiated is enough for PQ}
A data-independent implementation $\mathcal{I}$ is linearizable with respect to a data-independent specification $S$, if and only if $\mathcal{I}_{\neq}$ is linearizable with respect to $S_{\neq}$.
\end{restatable}