### 2017-04-20 co-regular-EPQ.tex

%!TEX root = draft.tex
\section{Reducing Linearizability of Priority Queues to Reachability}
\label{sec:co-regular of extended priority queues}

We show that the set of executions for which some projection fails the test $\textit{Check-PQ-Conc-NonRec}$ can be characterized using a set of register automata, modulo a value renaming. The possibility of renaming values (which is complete for checking data independent implementations) allows to simplify the reasoning about projections. Thus, we assume that all the operations which are not in the projection failing this test use the same distinguished value $\top$, different from those used in the projection. Then, it is enough to find an automata characterization for the set of executions $e$ for which $\textit{Check-PQ-Conc-NonRec}$ fails, or equivalently, for which one 
%a projection $e'$ fails $\textit{Check-PQ-Conc-NonRec}$ when one 
of the following three formulas is false:
\begin{align*}
\hspace{-5mm}
\Gamma(e) := \mathsf{Has\text{-}}\Gamma(e) \Rightarrow \exists \alpha.\ \Gamma\mathsf{\text{-}Conc}(e,\alpha)\mbox{ with $\Gamma\in \{\mathsf{EmptyRemove}$, $\mathsf{UnmatchedMaxPriority}$, $\mathsf{MatchedMaxPriority}\}$}
\end{align*}
Extending the automaton corresponding to some $\Gamma$ with self-loops that allow any operation with parameter $\top$ results in an automaton satisfying the following property called \emph{$\Gamma$-completeness}.

\begin{definition}
For $\Gamma\in \{\mathsf{EmptyRemove}$, $\mathsf{UnmatchedMaxPriority}$, $\mathsf{MatchedMaxPriority}\}$, an automaton $A$ is called \emph{$\Gamma$-complete} when for each data-independent implementation $\mathcal{I}$:

$A \cap \mathcal{I} \neq \emptyset$ if and only if there exists $ e \in \mathcal{I}$ and $e' \in \textit{proj}(e)$ such that $\Gamma(e')$ doesn't hold.
\end{definition}

The following result shows that deriving $\Gamma$-complete automata for each $\Gamma$ enables an effective reduction of checking linearizability of concurrent priority queue implementations to state reachability. It is a direct consequence of the above definitions.

\begin{restatable}{theorem}{ReduceEPQIntoStateReachability}
\label{lemma:reduce EPQ into state reachability}
Let $\mathcal{I}$ be a data-independent implementation, and $A(\Gamma)$ a $\Gamma$-complete automaton for each $\Gamma$. Then,
$\mathcal{I} \sqsubseteq \seqPQ$ if and only if $\mathcal{I} \cap A(\Gamma) = \emptyset$ for all $\Gamma$.
\end{restatable}

Section~\ref{ssec:aut} describes a $\mathsf{MatchedMaxPriority}$-complete automaton (TODO the other automata are discussed in Appendix~\ref{} or Icalp??) while Section~\ref{} discusses decidability results implied by the theorem above.

W.l.o.g., we assume that every implementation $\mathcal{I}$ behaves correctly, i.e., as a FIFO queue, when only values with the same priority are observed. More precisely, we assume that for every execution $e\in\mathcal{I}$ and every priority $p\in\mathbb{P}$, the projection of $e$ to values with priority $p$ is linearizable. This property can be checked separately using the techniques introduced for regular FIFO queues in~\cite{DBLP:conf/icalp/BouajjaniEEH15} (see Appendix~\ref{} for more details). Note that this assumption excludes some obvious violations, such as a $\textit{rm}(a)$ operation happens before a $\textit{put}(a,p)$ operation, for some $p$.

Also, we consider $\Gamma$-complete automata for $\Gamma\in \{\mathsf{UnmatchedMaxPriority}, \mathsf{MatchedMaxPriority}\}$, recognizing executions where any two occurring priorities are comparable (w.r.t. $\prec$). This is possible because any data-differentiated execution for which $\Gamma(e)$ is false has such a projection. Formally, given a data-differentiated execution $e$ and $p$ a maximal priority in $e$, $e\vert_{\preceq p}$ is the projection of $e$ to the set of values with priorities comparable to $p$. Then,

\begin{restatable}{lemma}{priExecutionIsEnough}
\label{lemma:pri execution is enough}
Let $\Gamma\in \{\mathsf{UnmatchedMaxPriority}, \mathsf{MatchedMaxPriority}\}$ and $e$ be a data-differentiated execution. Then, $\Gamma(e)$ is false if and only if $\Gamma(e\vert_{\preceq p})$ is false for some maximal priority $p$ in $e$.
\end{restatable}
\begin {proof} (Sketch)
TODO UPDATE THE FOLLOWING
To prove the $\textit{only if}$ direction, given $e \sqsubseteq l = u \cdot \textit{put}(x,p) \cdot v \cdot \textit{rm}(x) \cdot w \in \mathsf{MatchedMaxPriority}\mathsf{\text{-}Seq}(s,\alpha)$. Since there is no restriction to $p$-incomparable operations in $u$, $v$ and $w$, we can safely erase all $p$-incomparable values. Similar case holds for $\Gamma=\mathsf{UnmatchedMaxPriority}$.

To prove the $\textit{if}$ direction, we use a lemma in Appendix, which said that if the projection of an execution into a set $O_1$ of operations has linearization $l_1$ and the projection of the execution into $O_2$, the remanning operations, has linearization $l_2$, then the execution has a linearization whose projection into $O_1$ is $l_1$ and projection into $O_2$ is $l_2$.

Given $e\vert_{\preceq p} \sqsubseteq l = u \cdot \textit{put}(x,p) \cdot v \cdot \textit{rm}(x) \cdot w \in \mathsf{MatchedMaxPriority}\mathsf{\text{-}Seq}(s,\alpha)$. By above lemma we know that $e$ has a linearization $l' = u' \cdot \textit{put}(x,p) \cdot v' \cdot \textit{rm}(x) \cdot w'$, such that the projection of $l'$ into $p$-incomparable values is l. We can see that $u$, $v$ and $w$ are sub-sequence of $u'$, $v'$ an $w'$, respectively. Since $\mathsf{MatchedMaxPriority}\mathsf{\text{-}Seq}(s,\alpha)$ does not have restriction to $p$-incomparable values, we can see that $l' \in \mathsf{MatchedMaxPriority}\mathsf{\text{-}Seq}(s,\alpha)$. \qed
\end {proof}

%When the data domain and priority domain are both fixed to be finite set, the register automata can be simulated with finite automata, and by Lemma \ref{lemma:reduce EPQ into state reachability}, checking linearizable w.r.t $\seqPQ$ is decidable.

%
%
% and 
%
%Let $A(\Gamma)$ be a $\Gamma$-complete automaton for each $\Gamma$. The following theorem 
%
%
%We will describe this automaton only in the case of $\Gamma=\mathsf{MatchedMaxPriority}$. The rest can be found in Appendix~\ref{}. These automata satisfy the following property 
%
%the main technical challenge is to show that each of the following sets of executions can be characterized in terms of automata (modulo renaming of values):
%
%
% define a set of register automata whose product with a given implementation $\mathcal{I}$ accepts a linearizability violation whenever $\mathcal{I}$ is not linearizable.
%
%are precise enough to recognize every non-linearizable implementation.
%
%In this section, we divide checking linearizability w.r.t $\seqPQ$ into checking linearizablity w.r.t $\Gamma\mathsf{\text{-}Seq}(s,\alpha)$, each of which can be done in polynomial-time. Due to data-independence, violations to each $\Gamma\mathsf{\text{-}Seq}(s,\alpha)$ can be captured by register automata. This reduces checking linearizability w.r.t $\seqPQ$ into state reachability problem.
%
%
%
%\subsection{Co-Regular}
%\label{subsec:definition of co-regular}
%

%Let $\Gamma\mathsf{\text{-}Condition}(e)$ be $\mathsf{HasEmptyRemoves}(e)$, $\neg \mathsf{HasEmptyRemoves}(e) \wedge \mathsf{HasUnmatchedMaxPriority}(e)$, and $\mathsf{HasMatchedMaxPriority}(e)$, for $\Gamma = \mathsf{EmptyRemove}, \mathsf{UnmatchedMaxPriority}, \mathsf{MatchedMaxPriority}$, respectively. $\textit{Check-PQ-Conc}$ inspires us that, when $\Gamma\mathsf{\text{-}Condition}(e)$ holds, there should exist some $\alpha$, such that  $\Gamma\mathsf{\text{-}Conc}(e,\alpha)$ holds. The following lemma states this formally. Let $\textit{proj}(e)$ be the set of projections of $e$ into set of values. When refer to $\textit{proj}(e)$, we implicitly assume that each $\textit{rm}(\textit{empty})$ in $e$ has a ghost argument that is unique. We say that execution $l$ is linearizable w.r.t $\Gamma$, if we can find a sequence $s$ where $e \sqsubseteq s$ and $\Gamma\mathsf{\text{-}Seq}(s,\alpha)$ holds for some $\alpha$.

%\begin{restatable}{lemma}{EPQasMultiInMRpriforHistory}
%\label{lemma:EPQ as multi in MRpri for history}
%Given a data-differentiated execution $e$, $e \sqsubseteq \seqPQ$, if and only if, $\forall e' \in \textit{proj}(e)$, $\forall \Gamma\in \{\mathsf{EmptyRemove}$, $\mathsf{UnmatchedMaxPriority}$ $,\mathsf{MatchedMaxPriority}\}$,  $\Gamma\mathsf{\text{-}Condition}(e)$ $\Rightarrow$ $e$ is linearizable w.r.t $\Gamma$. 
%\end{restatable}

%Note that due to data-independence, register automata are used to detect some renamed violations, instead of store and checking directly. Let us introduce the notion of co-regular, which reduce checking linearizable w.r.t $\Gamma\mathsf{\text{-}Conc}(e,\alpha)$ into checking set of witness automata.
%
%%\vspace{-6pt}
%\begin{definition}\label{def:co-regular of rules of extended priority queues}
%Given $\Gamma\in \{\mathsf{EmptyRemove}$, $\mathsf{UnmatchedMaxPriority}$ $,\mathsf{MatchedMaxPriority}\}$. $\Gamma$ is co-regular, if there is a finite set $\textit{Auts}_{\Gamma}$ of register automata, such that for each data-independent implementation $\mathcal{I}$, we have that:
%
%$\textit{Auts}_{\Gamma} \cap \mathcal{I} \neq \emptyset$ if and only if $\exists e \in \mathcal{I}_{\neq},e' \in \textit{proj}(e)$, such that $\Gamma\mathsf{\text{-}Condition}(e')$ holds but $e$ does not linearizable w.r.t $\Gamma$. 
%
%$\seqPQ$ is co-regular, if each $\Gamma$ is co-regular.
%\end{definition}

\subsection{A $\mathsf{MatchedMaxPriority}$-complete automaton}\label{ssec:aut}

The formula $\mathsf{MatchedMaxPriority}$ is false for some execution $e$ when all the values of maximal priorities in $e$ are removed and there exists a value $x$ of maximal priority $p$ s.t. $\mathsf{MatchedMaxPriority\text{-}Conc}(e,x)$ is false. We consider two cases depending on whether $x$ is the only value with priority $p$. We denote by $\mathsf{MatchedMaxPriority}^{>}$ the strengthening of $\mathsf{MatchedMaxPriority}$ with this condition (i.e., all the values other than $x$ have a priority strictly smaller than $p$) and $\mathsf{MatchedMaxPriority}^{=}$ for the strengthening of the same formula with the negation of this condition (i.e., there exists another value $y$ with priority $p$).

Register automata is a kind of finite automata with registers used to deal with infinite data domain, where register can be assigned values and do simple operations. In this paper we use a restricted kind of register automata, which contains only one register, whose value is obtained once by guessing and never changed. Our register machine contains a register $r$. Its transition labels are $r=*,\textit{call}(\textit{put},d,g), \textit{ret}(\textit{put},d,g),$ $\textit{call}(\textit{rm},d),\textit{ret}(\textit{rm},d)$, where $d \in \mathbb{D} \cup \{ \textit{empty} \}$, and $g$ and $\top$ are guards. The first transition from initial state is a $r=*$ transition, and is used to assign an arbitrary value to $r$. A guard is a predicate chosen from $\{=r,< r,\top \}$, which represents equals the value of $r$, less than the value of $r$, and always holds, respectively. Given an execution $e = \alpha_1 \cdots \alpha_k$ of priority queue, register automata $\mathcal{A}$ accepts $e$, if there exists value $d_r\in \mathbb{D}$ and transitions $q_0 \xrightarrow{\beta_1} q_1 \ldots \xrightarrow{\beta_k} q_k$ of $\mathcal{A}$ from a initial state to a final state, such that for each $i$

\begin{itemize}
\setlength{\itemsep}{0.5pt}
\item[-] If $\alpha_i = \textit{call}(\textit{put},a,p)$, then $\beta_i= \textit{call}(\textit{put},a,g)$ and $g$ holds with $p$ and $d_r$. Similarly for $\textit{ret}(\textit{put},a,p)$.

\item[-] Else, $\alpha_i = \beta_i$.
\end{itemize}

\subsubsection{A $\mathsf{MatchedMaxPriority}^>$-complete automaton}
\label{subsec:co-regular of EPQ1Lar}

%We use two point to simply our proof of co-regular. The first point is to use the results in \cite{Bouajjani:2015} to ensure FIFO (first in first out) property of each single-priority projections. A execution where only values of one priority is putted is called a single-priority execution. It is obvious that each single-priority projection of sequences in $\seqPQ$ satisfy FIFO property. Let $\textit{toQueue}(e)$ be an execution generated from $e$ by transforming $\textit{put}$ and $\textit{rm}$ into $\textit{enq}$ and $\textit{deq}$, respectively, and then discarding priorities. In Appendix \ref{sec:appendix proof and definition in section definition of co-regular}, according to the result of queue in \cite{Bouajjani:2015}, we construct a set $\textit{Auts}_{\textit{sinPri}}$ of witness automata, and shows that they are enough to ensure that for each data-differentiated executions, each of its single-priority projection without $\textit{rm}(\textit{empty})$ to have ``FIFO'' property, as shown by the following lemma.
%
%\begin{restatable}{lemma}{AutoForEPQwithSignlePri}
%\label{lemma:automata for extended priority queue with single priority}
%
%Given a data-independent implementations $\mathcal{I}$ of extended priority queue, $\mathcal{I} \cap \textit{Auts}_{\textit{sinPri}} \neq \emptyset$, if and only if there exists $e \in \mathcal{I}_{\neq}$, $e' \in \textit{proj}(e)$, such that $e'$ is single-priority  without $\textit{rm}(\textit{empty})$, and $\textit{toQueue}(e')$ does not linearizable to queue.
%\end{restatable}
%
%According to Lemma \ref{lemma:automata for extended priority queue with single priority}, from now on, it is safe to assume that, for each data-differentiated execution, any of its single-priority projection without $\textit{rm}(\textit{empty})$ has ``FIFO'' property. This effectively exclude some obvious violations, such as $\textit{rm}(a)$ happens before $\textit{put}(a,\_)$.

%Lemma \ref{lemma:pri execution is enough} makes our co-regular proof more clear. One consequence of it is that for given $e\vert_{\preceq p}$, $\Gamma\mathsf{\text{-}Condition}(e')$ holds only for one $\Gamma$. 
%
%
%
%
%
%
%\subsection{Co-Regular of $\mathsf{MatchedMaxPriority}^{>}$}
%\label{subsec:co-regular of EPQ1Lar}

To facilitate the proof of co-Regular of $\mathsf{MatchedMaxPriority}$, we separate $\mathsf{MatchedMaxPriority\text{-}Seq}(e,x)$ into the following two sub-predicates, where the first one represents that $x$ is the only value with priority $p$ in $e$ , and the second one represents that there are more than one value with priority $p$ in $e$. Here $\textit{occurOnce}(p,l)$ holds if there is only one value with priority $p$ in $l$.

{\small
\begin{align*}
\mathsf{MatchedMaxPriority}^{>}\mathsf{\text{-}Seq}(e,x)=\mathsf{true} & \mbox{ iff  $\mathsf{MatchedMaxPriority}\mathsf{\text{-}Seq}(e,x)=\mathsf{true}$,} \\
&\hspace{4mm}\mbox{$\textit{occurOnce}(p,e)$ where $p$ is the priority of $x$} \\
\mathsf{MatchedMaxPriority}^{=}\mathsf{\text{-}Seq}(e,x)=\mathsf{true} & \mbox{ iff  $\mathsf{MatchedMaxPriority}\mathsf{\text{-}Seq}(e,x)=\mathsf{true}$,} \\
&\hspace{4mm}\mbox{$\neg\textit{occurOnce}(p,e)$ where $p$ is the priority of $x$} \\
\end{align*}
}

$\mathsf{MatchedMaxPriority}^{>}\mathsf{\text{-}Conc}(e,x)$ is obtained from $\mathsf{MatchedMaxPriority}\mathsf{\text{-}Conc}(e,x)$ by using $\mathsf{MatchedMaxPriority}^{>}\mathsf{\text{-}Seq}(e,x)$ instead of $\mathsf{MatchedMaxPriority\text{-}Seq}(e,x)$. And similarly for $\mathsf{MatchedMaxPriority}^{=}\mathsf{\text{-}Conc}(e,x)$.

We can also define two conditions, where $\textit{unmatched-priorities}(l)$ is the set of priorities occurring in $\textit{put}$ operations of $e$ for which there is $\textit{rm}$ operation removing the same value:

{\small
\begin{align*}
\mathsf{MatchedMaxPriority}^{>}\mathsf{\text{-}Condition}(e)=\mathsf{true} & \mbox{ iff $p\in \textit{matched-priorities}(e)$ for a maximal priority} \\
&\hspace{4mm}\mbox{$p\in priorities(e)$, $\textit{occurOnce}(p,e)$} \\
\mathsf{MatchedMaxPriority}^{=}\mathsf{\text{-}Condition}(e)=\mathsf{true} & \mbox{ iff $p\in \textit{matched-priorities}(e)$ for a maximal priority} \\
&\hspace{4mm}\mbox{$p\in priorities(e)$, $\neg\textit{occurOnce}(p,e)$} 
\end{align*}}

In this subsection, we briefly introduce the idea for proving co-regular of $\mathsf{MatchedMaxPriority}^{>}$. The proof of this subsection can be found in Appendix \ref{sec:appendix proof and definition in section co-regular of EPQ1Lar}. The notion of left-right constraint used for priority queue is inspired by left-right constraint of queue \cite{Bouajjani:2015}.

Given a data-differentiated $\_$-execution $e$ such that $\mathsf{MatchedMaxPriority}^{>}\mathsf{\text{-}Condition}(e)$ holds, Lemma \ref{lemma:automata for extended priority queue with single priority} is not enough for ensure linearizable w.r.t $\mathsf{MatchedMaxPriority}^{>}$. This is because that it is possible that $\mathsf{MatchedMaxPriority}^{>}\mathsf{\text{-}Conc}(e,x)$ does not holds because of interaction between actions of multiple priorities. One such example is shown in \figurename~\ref{fig:introduce gap for EPQ1Lar}. We call the time interval from $\textit{ret}(\textit{put},x)$ to $\textit{cal}(rm,x)$, or from $\textit{ret}(\textit{put},x)$ when $\textit{cal}(rm,x)$ does not exist, the interval of value $x$. In \figurename~\ref{fig:introduce gap for EPQ1Lar}, we draw the interval of each value by dashed line. Here we assume that $p_1 \prec p_4$, $p_2 \prec p_4$ and $p_3 \prec p_4$. We can not find a sequence $s$ where $e \sqsubseteq s$ and $\mathsf{MatchedMaxPriority}^{>}\mathsf{\text{-}Seq}(s,b)$ holds, since each time point from $\textit{cal}(\textit{rm},b)$ to $\textit{ret}(\textit{rm},b)$ is in interval of some value with smaller priority, and $\textit{rm}(b)$ can not happen in interval of value with smaller priority.

\begin{figure}[htbp]
  \centering
  \includegraphics[width=0.4 \textwidth]{figures/PIC-HIS-INTRO-GAP-EPQ1L.pdf}
%\vspace{-10pt}
  \caption{An execution that does not linearizable w.r.t $\mathsf{MatchedMaxPriority}^{>}$}
  \label{fig:introduce gap for EPQ1Lar}
\end{figure}

In Appendix \ref{sec:appendix proof and definition in section co-regular of EPQ1Lar}, we introduce the notion of left-right constraint, which is a graph, and the existence of cycle though value with maximal priority in it formally modelled the case in \figurename~\ref{fig:introduce gap for EPQ1Lar}. For example, in \figurename~\ref{fig:introduce gap for EPQ1Lar}, $b$ is covered by $f$, $e$ and $d$. We then prove that getting rid of cycle though value with maximal priority in left-right constraint is enough for ensure linearizable w.r.t $\mathsf{MatchedMaxPriority}^{>}$.

When there is a cycle $d_1 \rightarrow \ldots \rightarrow d_m \rightarrow x \rightarrow d_1$ through value $x$ with maximal priority in left-right constraint, we say that $x$ is covered by $d_1,\ldots,d_m$. Assume that $x$ is covered by $d_1,\ldots,d_m$, we can safely rename $x$ into $b$, rename $d_1,\ldots,d_m$ into $a$ and rename all other value into $c$ by data-independence. Such execution can be recognized by witness automata, since between the first $\textit{ret}(\textit{put},a,\_)$ and the last $\textit{call}(\textit{rm},a)$ (if exists), $\textit{ret}(\textit{put},a,\_)$ and $\textit{call}(\textit{rm},a)$ occurs in pair and there is no need for count. In Appendix \ref{sec:appendix proof and definition in section co-regular of EPQ1Lar}, we construct a set $\textit{Auts}_{\textit{1-lar}}$ of register automata, and use $\textit{Auts}_{\textit{1-lar}}$ to prove that $\mathsf{MatchedMaxPriority}^{>}$ is co-regular, as stated by the following lemma.

\begin{restatable}{lemma}{EPQOneLarisCoRegular}
\label{lemma:EPQ1Lar is co-regular}
$\mathsf{MatchedMaxPriority}^{>}$ is co-regular.
\end{restatable}

\subsection{A $\mathsf{MatchedMaxPriority}^=$-complete automaton}
\label{subsec:co-regular of EPQ1Equal}

In this subsection, we briefly introduce the idea for proving co-regular of $\mathsf{MatchedMaxPriority}^{=}$. The proof of this subsection can be found in Appendix \ref{sec:appendix proof and definition in section co-regular of EPQ1Equal}.

Given a data-differentiated $\_$-execution $e$ such that $\mathsf{MatchedMaxPriority}^{=}\mathsf{\text{-}Condition}(e)$ holds, Lemma \ref{lemma:automata for extended priority queue with single priority} and $\textit{Auts}_{\textit{1-lar}}$ are sill not enough for ensure that $e$ linearizable w.r.t $\mathsf{MatchedMaxPriority}^{=}$. This is because that given values $a$ and $b$ with maximal priority, it is possible that all the possible linearization points of $\textit{rm}(b)$ are disabled by $\textit{rm}(b)$. We give an example of such execution $e$ in \figurename~\ref{fig:introduce pb order}. The execution $e$ of \figurename~\ref{fig:introduce pb order} is not linearizable w.r.t $\mathsf{MatchedMaxPriority}^{=}$, even if $h \vert_{p_1}$ and $h \vert_{p_4}$ both linearizable w.r.t $\mathsf{MatchedMaxPriority}^{>}$, and either $a$ or $b$ is covered by values with smaller priority. The explanation of $e$ not linearizable w.r.t $\mathsf{MatchedMaxPriority}^{=}$ is as follows: Since $\textit{put}(a,p_4) <_{\textit{hb}} \textit{put}(b,p_4)$, the linearization points of $\textit{rm}(a)$ should before the linearization point of $\textit{rm}(b)$. According to the previous section, which says that linearizaton point of larger priority can be happen in the interval of values with smaller priority. Thus, the time intervals identified with dotted lines in \figurename~\ref{fig:introduce pb order} are are possible position to locate the linearization point of $\textit{rm}(b)$. However, each of them is before $\textit{call}(\textit{rm},a)$.

\begin{figure}[htbp]
  \centering
  \includegraphics[width=0.6 \textwidth]{figures/PIC-HIS-INTRO-PB-ORDER-EPQ.pdf}
%\vspace{-10pt}
  \caption{An execution that does not linearizable w.r.t $\textit{MS}(\textit{EPQ}_1^{=})$}
  \label{fig:introduce pb order}
\end{figure}

One point need to be modeled in \figurename~\ref{fig:introduce pb order} is that ``a is putted before b''. Let us introduce put-before order $<_{\textit{pb}}$ to formally state that ``an value is putted before another value''. Given a data-differentiated $\_$-execution $e$ and two values $a,b$ with maximal priority, $a <_{\textit{pb}} b$, if one of the following cases holds: (1) $\textit{put}(a,\_) <_{\textit{hb}} \textit{put}(b,\_)$, (2) $\textit{rm}(a) <_{\textit{hb}} \textit{rm}(b)$, or (3) $\textit{rm}(a) <_{\textit{hb}} \textit{put}(b,\_)$. Sometimes we use $a <_{\textit{pb}}^A b$, $a <_{\textit{pb}}^B b$ and $a <_{\textit{pb}}^C b$ to explicitly distinguish above three cases. Let $<_{\textit{pb}}^*$ be the transitive closure of $<_{\textit{pb}}$. Intuitively, $a <_{\textit{pb}}^* b$ means that we can infer that $a$ is putted before $b$ with the help of several intermediate values.

Another point need to be modeled in \figurename~\ref{fig:introduce pb order} is the possible time points of $\textit{rm}(b)$, which is formally modeled by gap-point.

\begin{definition}\label{def:gap-point for matched put and rm operations}
Given a data-differentiated $p$-execution $e$ and two method events $\textit{put}(x,p),\textit{rm}(x)$ of $e$. We say that a time-point $o$ is a gap-point of $x$, if $o$ is after $\textit{call}(\textit{put},x,p)$ and $\textit{call}(\textit{rm},x)$, is before $\textit{ret}(\textit{rm},x)$, and is not in interval of any values with smaller priority.
\end{definition}

The case of \figurename~\ref{fig:introduce pb order} can be formally described as follows: $a <_{\textit{pb}}^* b$, while the right-most gap-point of $b$ is before $\textit{call}(\textit{rm},a)$ or $\textit{call}(\textit{put},a,p_4)$. The following lemma states that getting rid of such case is enough for ensure linearizable w.r.t $\mathsf{MatchedMaxPriority}^{=}$.

\begin{restatable}{lemma}{EPQOneEqualAsPBandGP}
\label{lemma:EPQ1Equal as pb order and gap-point}
Given a data-differentiated $p$-execution $e$ where $\mathsf{MatchedMaxPriority}^{=}\mathsf{\text{-}Condition}(e)$ holds. $e$ is not linearizable w.r.t $\mathsf{MatchedMaxPriority}^{=}$, if and only if there exists $x$ and $y$ with maximal priority $p$, such that $y <_{\textit{pb}}^* x$, and the rightmost gap-point of $x$ is before $\textit{call}(\textit{put},y,p)$ or $\textit{call}(\textit{rm},y)$ in $e$. 
\end{restatable}

\begin {proof} (Sketch)

We have already intuitively explain the proof of the $\textit{if}$ direction. To prove the $\textit{only if}$ direction, we prove its contrapositive. %, where we already know that $\forall x, y$ with maximal priority, if $y <_{\textit{pb}}^* x$, then the rightmost gap-point of $x$ is after $\textit{cal}(\textit{put},y,\textit{pri})$ and $\textit{cal}(\textit{rm},x)$. We need to prove that $e \sqsubseteq \textit{MS}(\textit{EPQ}_1^{=})$. %Or we can say, we need to explicitly construct linearization of $e$.

Let $\textit{Items}(e,p)$ be the set of values with priority $p$ in $e$. We introduce another lemma (Lemma \ref{lemma:maximal in pb and gap-point make a candidate of EPQ1Equal} in Appendix), which states that if we can find a value $\exists g_1 \in \textit{Items}(e,p)$, such that $\forall g_2 \in \textit{Items}(e,p)$, (1) $g_1 \nless_{\textit{pb}} g_2$, and (2) the right-most gap-point of $g_1$ is after $\textit{call}(\textit{put},g_2,p)$ and $\textit{call}(\textit{rm},g_2)$. Then $e$ is linearizable w.r.t $\mathsf{MatchedMaxPriority}^{=}$. 
The remaining of our proof is a loop for searching such $g_1$ in $e$. Let $e_p$ be the projection of $e$ into values of priority $p$, and let $l_p$ be the linearization of $e$ which satisfy the requirements of queue. Let the initial value of $a_1$ be the last inserted value in $l_p$.

\begin{itemize}
\setlength{\itemsep}{0.5pt}
\item[-] check if $a_1$ satisfy the conditions of Lemma \ref{lemma:maximal in pb and gap-point make a candidate of EPQ1Equal}. If it does, then this loop terminates.

\item[-] Otherwise, $\exists a_2 \in \textit{Items}(e,p)$, such that the rightmost gap-point of $a_1$ is before $\textit{cal}(\textit{put},a_2,p)$ or $\textit{cal}(\textit{rm},a_2)$ in $e$. By assumption, we know that $a_2 \nless_{\textit{pb}} a_1$, and each gap-point of $a_2$ is after the rightmost gap-point of $a_1$.

    \begin{itemize}
    \setlength{\itemsep}{0.5pt}
    \item[-] If $\forall b \in \textit{Items}(e,p)$, $a_2\not<_{\textit{pb}}b$. Then we go to next round of the loop and make $a_2$ to be the ``new-round $a_1$''.
    \item[-] Otherwise, there exists $a_3 \in \textit{Items}(e,p)$ such that $a_2 <_{\textit{pb}}^* a_3$. It is not hard to see that there is no cycle in $<_{\textit{pb}}$, and it is safe to assume that $a_3$ is maximal w.r.t $<_{\textit{pb}}^*$.

        By assumption,we know that the rightmost gap-point of $a_3$ is after $\textit{call}(\textit{put},a_2,p)$ and $\textit{call}(\textit{rm},a_2)$. Therefore, we can see that the rightmost gap-point of $a_3$ is after the rightmost gap-point of $a_1$. Then we go to next round of the loop and make $a_3$ to be the ``new-round $a_1$''.
    \end{itemize}
\end{itemize}

Let $a^i$ be the $a_1$ in the $\textit{i-th}$ round of loop. We can see that the rightmost gap-point of $a^j$ is after the rightmost gap-point of $a^i$ for each $i<j$. Therefore, the loop finally terminates at some $a^f$ that satisfies the condition of Lemma \ref{lemma:maximal in pb and gap-point make a candidate of EPQ1Equal}. This implies that $e$ is linearizable w.r.t $\mathsf{MatchedMaxPriority}^{=}$ and completes the proof of the $\textit{only if}$ direction. \qed
\end {proof}

The result of Lemma \ref{lemma:EPQ1Equal as pb order and gap-point} still needs to be impove, since obtaining $y <_{\textit{pb}}^* x$ may requires arbitrary many intermediate values, and it is hard to store unbounded values by automata. Fortunately, we prove that the number of intermediate values is essentially bounded, as stated by the following lemma.

\begin{restatable}{lemma}{OBOrderHasBoundedLength}
\label{lemma:ob order has bounded length}
Given a data-differentiated execution $e$. Assume that $a <_{\textit{pb}} a_1 <_{\textit{pb}} \ldots <_{\textit{pb}} a_m <_{\textit{pb}} b$, then one of the following cases holds:

\begin{itemize}
\setlength{\itemsep}{0.5pt}
\item[-] $a <_{\textit{pb}}^A b$, $a <_{\textit{pb}}^B b$ or $a <_{\textit{pb}}^C b$,

\item[-] $a <_{\textit{pb}}^A a_i <_{\textit{pb}}^B b$, or $a <_{\textit{pb}}^B a_i <_{\textit{pb}}^A b$, for some $i$.
%\item[-] $a <_{\textit{pb}}^B a_i <_{\textit{pb}}^A a_j <_{\textit{pb}}^B b$, for some $i$ and $j$,
\end{itemize}
\end{restatable}

By Lemma \ref{lemma:ob order has bounded length}, we can use register automata to detect $a <_{\textit{pb}}^* b$ by enumerating all possible enumerations of operations of $a$, $b$ and possibly $a_1$. Since some combinations of $<_{\textit{pb}}^*$ and gap-points are conflict, we finally reduce the number of potential enumerations of operations of $a$, $b$ and possibly $a_1$ into only five in Appendix \ref{sec:appendix proof and definition in section co-regular of EPQ1Equal}.

In Appendix \ref{sec:appendix proof and definition in section co-regular of EPQ1Equal}, we construct a set $\textit{Auts}_{\textit{1-eq}}$ of register automata, and prove that $\mathsf{MatchedMaxPriority}^{=}$ is co-regular, as stated by the following lemma.

\begin{restatable}{lemma}{EPQOneEqualIsCoRegular}
\label{lemma:EPQ1Equal is co-regular}
$\mathsf{MatchedMaxPriority}^{=}$ is co-regular.
\end{restatable}

One such enumeration and its register automata is shown in \figurename~\ref{fig:an enumeration and its witness automaton}. $o$ is the rightmost gap-point of $b$, we rename the values that ``covers the time interval from $\textit{call}(\textit{rm},a)$ to $\textit{ret}(\textit{rm},b)$'' (see Appendix \ref{sec:appendix proof and definition in section co-regular of EPQ1Equal}) into $d$, and rename the renaming values into $e$. In this figure, $c = \textit{call}(\textit{put},e,\top),\textit{ret}(\textit{put},e,\top)$, $\textit{call}(\textit{rm},e), \textit{ret}(\textit{rm},e),\textit{call}(\textit{rm},\textit{empty}),\textit{ret}(\textit{rm},\textit{empty})$, $c_1 = c + \textit{call}(\textit{put},d,<r)$, $c_2 = c_1 + \textit{ret}(\textit{put},b,r)$, $c_3 = c_2 + \textit{ret}(\textit{rm},d)$, $c_4 = c + \textit{ret}(\textit{put},b,r) + \textit{ret}(\textit{rm},d)$.

\begin{figure}[htbp]
  \centering
  \includegraphics[width=0.8 \textwidth]{figures/PIC-Enumeration-WitnessAutomata.pdf}
%\vspace{-10pt}
  \caption{An enumerations and its register automaton}
  \label{fig:an enumeration and its witness automaton}
\end{figure}

\subsection{Reduce into State Reachability}
\label{subsec:combine step-by-step linearizability and co-regular}

We also prove that $\mathsf{UnmatchedMaxPriority}$ are co-regular in Appendix \ref{subsec:appendix co-regular of EPQ2Lar}, Appendix \ref{subsec:appendix co-regular of EPQ2Equal}, and prove that $\mathsf{EmptyRemove}$ are co-regular in Appendix \ref{subsec:co-regular of EPQ3}. Then we can see that $\seqPQ$ is co-regular.

\begin{restatable}{lemma}{EPQIsCoRegular}
\label{lemma:EPQ is co-regular}
$\seqPQ$ is co-regular.
\end{restatable}

By Lemma \ref{lemma:EPQ as multi in MRpri for history} and Lemma \ref{lemma:EPQ is co-regular}, we can finally reduce the verification of linearizability w.r.t $\seqPQ$ into a reachability problem, as illustrated by the following theorem, where $\textit{Auts}_{\textit{EPQ}}$ is the set of witness automata of this section.

\begin{restatable}{theorem}{ReduceEPQIntoStateReachability}
\label{lemma:reduce EPQ into state reachability}
Given a data-independence implementation $\mathcal{I}$. $\mathcal{I} \sqsubseteq \seqPQ$, if and only if, $\mathcal{I} \cap \textit{Auts}_{\textit{EPQ}} = \emptyset$.
\end{restatable}

When the data domain and priority domain are both fixed to be finite set, the register automata can be simulated with finite automata, and by Lemma \ref{lemma:reduce EPQ into state reachability}, checking linearizable w.r.t $\seqPQ$ is decidable.